---
title: "had.27.46a20: Reproduce the advice for WGMIXFISH"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, settings, include=FALSE,echo=FALSE}
rm(list=ls())

knitr::opts_chunk$set(
  fig.width = 9, message=FALSE, warning=FALSE,eval=TRUE,include=TRUE,echo=TRUE
)
options(digits=4)

```

# Setup
```{r, setup,results='hide'}

library(knitr)
library(FLCore)
library(FLasher)
library(mixfishtools)
library(tidyr)
library(ggplot2)
library(ggplotFL)
library(kableExtra)

load("data/init.RData") # Contains useful stuff such as reference points

# forecast year definitions
yrAssess <- ay-1 # final year of assessment data
yrNow <- ay # intermediate year
yrTAC <- ay+1 # advice year
yrTACp1 <- ay+2 # advice year +1 (needed to get SSB at end of yrTAC)

```


# Gather advice sheet values

A TAC constraint was not used this year. The intermediate year assumption was F *status quo*.

Most of these values should be as reported on the advice sheet. However, for some the correct values for comparison are reported in *Assessment summary and forecast results\_ NoSh haddock\_ WGNSSK_2025.xslx*. These are:

The SSB values in `r ay` as the value in the spreadsheet uses the stock weights assumed for the forecast which are different to those used in the assessment.

The catch and landings total in `r ay-1` as the value in the spreadsheet are those as estimated by SAM for `r ay-1` rather than the reported InterCatch value.

```{r, advice-sheet-values}

# Gather forecast info from advice sheet ------------------------------------------

rec_assum <- rep(2014079,3) # take from table 1 in advice sheet

Fsq <- 0.06 # from table 1 in advice sheet
TAC_cont <- F

# get these values from assessment and forecast summary spreadsheet, not advice sheet.
SSB_ay <- 740576  # This is because SSB gets updated due to forecast weights.
catch_ay_1 <- 60470 # We want SAM estimate of catches not InterCatch value.
landings_ay_1 <- 44771 #  We want SAM estimate of landings not InterCatch value.

# Reported output from single stock headline advice. See above for estimated values of landings and catch
# yrAssess values (data yr) - take from table 9 in advice sheet
# yrNow values (int yr) - take from table 1 in advice sheet. SSB isn't on advice sheet - see SSB_ay.
# yrTAC values (advice year) - take from table 2, SSB comes from table 1
# yrTACp1 values (SSB only) - take from table 2

stfRef <- data.frame(
  model = "NSSK",
  year = yrAssess:yrTACp1,
  catch = c(catch_ay_1, 54563, 108301, NA),
  landings = c(landings_ay_1, 43332, 91316, NA),
  fbar = c(0.06, 0.06, 0.167, NA),
  ssb = c(771472, SSB_ay, 607965, 470142)
)

```

# Get assessment results

```{r, assess-res}
# load assessment results ------------------------------------

load("output/had_27_46a20_FLStock object_model estimates_2025.Rdata")
load("output/had_27_46a20_FLStock object_input data_2025.Rdata")

stkInp <- stock.data
stkEst <- window(stock,end=yrAssess)

units(stkInp@harvest) <- "f"

L <- FLStocks(list(input = stkInp, estimated = stkEst))
plot(L) + 
  aes(linetype = stock) +
  scale_color_manual(values = c(8,1)) + 
  scale_linetype_manual(values = c(1,2)) +
  theme_bw()

```

# Forecast setup

The mean weights at age for the stock, catch, landings and discards get updated here to externally modelled values (cohort model).

```{r, forecast-setup}
# Extend FLStock, add forecast info ---------------------------------------

# extend FLStock object
stkProj <- stf(object = stkEst, nyears = 3, wts.nyears = 3, 
               fbar.nyears = 3, f.rescale = TRUE, disc.nyears = 3)

# stock-recruitment model (manual input within a geometric mean model)
srPar <- FLPar(rec_assum, 
               dimnames = list(params="a", year = c(yrNow, yrTAC, yrTACp1), iter = 1))

srMod <- FLSR(model = "geomean", params = srPar)


# replace individual weights
ca.frct.wt <- read.table("output/Forecast/had.27.46a20 - Forecast weights - catch-at-age.txt")
st.frct.wt <- read.table("output/Forecast/had.27.46a20 - Forecast weights - stock-at-age.txt")
lan.frct.wt <- read.table("output/Forecast/had.27.46a20 - Forecast weights - landings-at-age.txt")
dis.frct.wt <- read.table("output/Forecast/had.27.46a20 - Forecast weights - discards-at-age incl BMS and IBC.txt")
colnames(ca.frct.wt) <-colnames(st.frct.wt) <-colnames(lan.frct.wt) <-colnames(dis.frct.wt) <- c(0:7,"8+")

# print out these forecast weights for transparency
catch.wt(stkProj)[, ac(yrNow:yrTACp1)][] <- t(ca.frct.wt) # catch weight-at-age
kable(ca.frct.wt, digits = 4,caption="Forecast weights for the catch")

landings.wt(stkProj)[, ac(yrNow:yrTACp1)][] <- t(lan.frct.wt) # landings weight-at-age
kable(lan.frct.wt, digits = 4,caption="Forecast weights for the landings")

discards.wt(stkProj)[, ac(yrNow:yrTACp1)][] <- t(dis.frct.wt) # discards weight-at-age
kable(dis.frct.wt, digits = 4,caption="Forecast weights for the discards")

stock.wt(stkProj)[, ac(yrNow:yrTACp1)][] <- t(st.frct.wt) # stock weight-at-age
kable(st.frct.wt, digits = 4,caption="Forecast weights for the stock")

# View the extended FLStock
df <- as.data.frame(stkProj)
df <- subset(df, slot %in% c("landings.wt", "discards.wt", "catch.wt", "m", "mat", "harvest") & year > (yrAssess-20))
df$forecast <- df$year %in% c(yrNow, yrTAC, yrTACp1)

ggplot(df) + aes(x = year, y = data, group = age, color = forecast) +
  facet_wrap(~slot, scales = "free_y") +
  geom_line(show.legend = F) + 
  scale_color_manual(values = c(8,1)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
  theme_bw()

```

# Set controls for the forecast

Fmsy is `r Fmsy`. A TAC constraint was not used for the intermediate year.

```{r, forecast-control}
# Make forecast control object --------------------------------

if(TAC_cont){ # TAC constraint used for intermediate year
  TACNow <- TAC
  
  ctrl <- fwdControl( 
    data.frame(
      year = c(yrNow, yrTAC, yrTACp1),
      value = c(TACNow, Fmsy, Fmsy),
      quant = c("catch", "f", "f")
    )
  )
  
}else{ #Fsq used for intermediate year
  ctrl <- fwdControl( 
    data.frame(
      year = c(yrNow, yrTAC, yrTACp1),
      value = c(1, Fmsy, Fmsy),
      quant = c("f"),
      relYear = c(yrAssess, NA, NA)                               
    )
  )
}

```

# Run the forecast

```{r, forecast}

# Run forecast ----------------------------------------------------

# projection
stkProj <- fwd(object = stkProj, control = ctrl, sr = srMod)

# plot
L <- FLStocks(list(assessment = stkEst, 
                   forecast = stkProj[,ac(yrAssess:yrTACp1)]))
plot(L) + 
  # aes(linetype = stock) +
  scale_color_manual(values = c(8,1)) + 
  # scale_linetype_manual(values = c(2,1)) +
  theme_bw()

df <- data.frame(year = yrAssess:yrTACp1, 
                 catch = c(catch(stkProj)[, ac(yrAssess:yrTACp1)]),
                 fbar = c(fbar(stkProj)[, ac(yrAssess:yrTACp1)])
)

kable(df, digits = 3)


```

# Compare forecast results

```{r, plot}


# WG2025 forecast n at age for 2025
# 0                    1                 2                  3                  4                  
# 1746746.2823940709  447459.9367602070  159355.9434177537  632360.9723833080  284049.7020637528 
# 5                  6                   7                  8 
# 630354.3028598678  287458.5550952689   23292.7518568312   16172.3850534831 

# Plot ------------------------------------------

# stfRef is defined in chunk: advice_sheet_values

stfDet <- data.frame(
  model = "FLR",
  year = ac(yrAssess:yrTACp1),
  catch = c(catch(stkProj[,ac(yrAssess:yrTACp1)])),
  landings = c(landings(stkProj[,ac(yrAssess:yrTACp1)])),
  fbar = c(fbar(stkProj[,ac(yrAssess:yrTACp1)])),
  ssb = c(ssb(stkProj[,ac(yrAssess:yrTACp1)]))
)  

df <- merge(stfRef, stfDet, all = T)
df <- pivot_longer(df, cols = c(catch, landings, fbar, ssb), 
                   names_to = "variable", values_to = "value")
df <- df |>
  filter(
    (variable %in% c("catch", "landings", "fbar") & year <= yrTAC) |
      (variable %in% c("ssb") & year <= yrTACp1))

df2 <- pivot_wider(df, names_from = model, values_from = value)
df2$percErr <- round((df2$FLR - df2$NSSK)/df2$NSSK * 100, 1)

ggplot(df) + aes(x = year, y = value, group = model, color = model, shape = model) +
  facet_wrap(~variable, scales = "free_y") +
  geom_line() +
  geom_point(size = 3, stroke = 1) +
  scale_shape_discrete(solid = F) +
  coord_cartesian(ylim = c(0, NA)) +
  theme_bw()

kable(df2, digits = 3)

ggplot(df2) + aes(x = year, y = percErr) + 
  facet_wrap(~variable) +
  geom_col(fill = 4, color = 4) +
  geom_hline(yintercept = 0, linetype = 1) +
  geom_hline(yintercept = c(-10,10), linetype = 3) + 
  theme_bw()

```
